Very simplified likelihood

### 2024-02-23 inst/ejs-sample.tex
work

\begin{equation}\label{likeforrn}
  \mathcal{L}_r(a)=\begin{cases}
\frac{b}{a+ n   b}\cdot
\frac{b}{a+(n-1)b}\cdot
\frac{b}{a+(n-2)b}\cdots
\frac{a}{a+(n-r)b}\qquad &{r<n}\\
\frac{b}{a+ n   b}\cdot
\frac{b}{a+(n-1)b}\cdot
\frac{b}{a+(n-2)b}\cdots
\frac{b}{a+b}\qquad &{r=n}
  \end{cases}
\end{equation}

Although these expressions do have closed-form analytical forms, it is
not helpful in numerical work because they tend to be very unstable,
especially where $kb\simeq n$ for some integer $k$, as limiting forms
have to be used.  Note carefully that the expressions in
equation~\ref{likeforrn} are likelihoods, not probabilities, and do
not sum to 1.

@@

Previous related work presented the case $n=2$ [in current notation].
Considering the case $n=7$ gives us $7+1=8$ distinct likelihood
functions, Figure~\ref{ninelikes}.  We see an unanticipated asymmetry
between $r$ and $n-r$, and in particular that
$\hat{a}\geqslant\frac{r}{n}$, with equality only in the edge-cases
$r=0$ or $n$.  For $r=3$ one might expect that $\hat{a}=\frac{1}{2}$
on the grounds that the focal competitor's performance was median.
However, $\hat{a}\simeq 0.514$.

>>>>

\begin{eqnarray}\label{likeforrn1}
  \mathcal{L}_{r,n}(a) &=& 
\frac{b}{a+ n   b}\cdot
\frac{b}{a+(n-1)b}\cdot
\frac{b}{a+(n-2)b}\cdots
\frac{a}{a+(n-r)b}\\ \label{likeforrn2}
&=& \frac{B-1}{(B+n-1)(B+n-2)\cdots(B+n-r-1)}\\ \label{likeforrn3}
&=& \frac{(B-1)(B+n-r-2)!}{(B+n-1)!}
\end{eqnarray}

\noindent where $B=1/b$.  Note carefully that these expressions are
likelihoods, not probabilities, and do not sum to 1.  Asymptotic
analysis of expression~\ref{likeforrn3} shows that

\begin{equation}
\log\mathcal{L}_{r,n}(a)=\log(B-1)-B\frac{r+1}{n} 
+ K + \mathcal{O}\left(n^{-2}\right)
\end{equation}

Using $\mathcal{S}=\log(B-1)-B\frac{r+1}{n}$ as an approximate support
thus incurs an error of only $\mathcal{O}(n^{-2})$.  Observing that
$\partial^2\mathcal{S'}/\partial B^2=-(B-1)^{-2}<0$, the evaluate
would be unique; we have
$\hat{B}=\frac{n+r+1}{r+1}+\mathcal{O}\left(n^{-1}\right)$, or
alternatively $\hat{a}=\frac{n}{n+r+1}$.  I present some numerical
results below.

Previous related work presented the case $n=2$ [in current notation].
Considering the case $n=7$ gives us $7+1=8$ distinct likelihood
functions, Figure~\ref{ninelikes}.  We see an unanticipated asymmetry
between $r$ and $n-r$.  For $r=3$ one might expect that
$\hat{a}=\frac{1}{2}$ on the grounds that the focal competitor's
performance was median.  However, $\hat{a}\simeq 0.64$.

@@

### 2024-02-23 inst/very_simplified_likelihood.Rmd
work

```{r,label=likevis,cache=F}
plotter <- function(...){
a <- seq(from=0,to=1,by=0.01)
n <- 7
jj <- f_vec_arn(a,3,n,log=FALSE)
plot(a,jj/max(jj,na.rm=TRUE),type='n')
grid()
for(r in seq(from=0,to=n)){
  y <- f_vec_a(a,r,n,log=FALSE)
  if(r==0){y[length(y)] <- 1}  
  y <- y/max(y,na.rm=TRUE)
  if(r>0){y[length(y)] <- 0}
  points(a,y,type='l',lwd=4,col=rainbow(n+1)[r+1])
}
}
plotter()
pdf(file="ninelikes.pdf")
plotter()
dev.off()
```

>>>>

```{r,label=likevis,cache=F}
plotter <- function(...){
a <- seq(from=0,to=1,by=0.01)
n <- 7
jj <- f_vec_arn(a,3,n,log=FALSE)
plot(a,jj/max(jj,na.rm=TRUE),type='n')
grid()
for(r in seq(from=0,to=n)){
  y <- f_vec_a(a,r,n,log=FALSE)
  if(r==0){y[length(y)] <- 1}  
  y <- y/max(y,na.rm=TRUE)
  if(r>0){y[length(y)] <- 0}
  points(a,y,type='l',lwd=4,col=rainbow(n+1)[r+1])
  abline(v=0.64)
}
}
plotter()
pdf(file="ninelikes.pdf")
plotter()
dev.off()
```
